---
title: "Age of Planes"
author: "Gabriel Guerrero"
date: "`r format(Sys.time(), '%B %d, %Y')`"
execute:
  keep-md: true
  warning: false
format:
  html:
    code-fold: true
    code-tools: true
---

## 1. Background

In a previous task we used the flights tibble from the nycflights13 package to answer questions about flight delays. That same package contains four other tibbles that are related to flights and can help us answer more complex questions. Run data(package = "nycflights13") to see a list of the tibbles with a brief description.

For this task we are trying to answer two questions:

1. What is the relationship between departure delays and age of the plane? 
2. What is the relationship between airline carrier and age of the plane?

## 2. Use a flights dataset

```{r}
#| label: load_packages
#| echo: true

# library tidyverse and nycflights13 installed
library(tidyverse)
library(nycflights13)
#install.packages('nycflights13')
#data(package = "nycflights13")

```

#### Example of the first rows of the data

```{r}
#| echo: false
str(flights, 5)
str(planes,5)

```


```{r}
#| label: join
#| echo: false
## Join by tail number
dfcom <- flights%>%
        left_join(planes,c("tailnum"="tailnum"))

#calculate the age of the plane that was used.
dfcom<-mutate(dfcom, age=year.x - year.y)
dfcom <-dfcom%>% 
  rename(creation_year = year.y, dept_year = year.x)

```

## Chart
Create a chart that looks at the distribution of departure delays by age (or age group). Do you notice any patterns?
Create a new column that groups age into 0-10, 11-20, 21-30, and >30

```{r}
#| label: chart1
#| echo: false
chart_1<- mutate(dfcom, age_references=case_when(
    age>= 0 & age <= 10 ~ "0-10",
    age>= 11 &age <= 20 ~ "11-20",
    age>= 21 & age <= 30 ~ "21-30",
    age> 30 ~ ">30",
    TRUE ~ "others"  
))

table <- select(chart_1, age,age_references,creation_year, dept_year, dep_delay)



sum <- table%>%
  group_by(age_references)%>%
 summarise(tot_delays = sum(dep_delay > 0, na.rm = TRUE)) 
    
ggplot(sum, aes(x = age_references, y = tot_delays, fill = age_references)) +
  theme_light()+
  geom_bar(stat = "identity")+
  labs(title = "Flights", y = "Delay Time", x = "Group Ages")+
  theme(plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        panel.grid.major.x=element_blank(),
        panel.grid.minor.y = element_blank()
        )+
scale_y_continuous(breaks =seq(0,50000, by=10000))

#head(table, 5)
```

## Create a chart that looks at the distribution of age by carrier.

```{r}
#| label: bycarrier
#| echo: false

join_carrier<- flights%>%
  left_join(planes, by = "tailnum") %>%
  left_join(airlines, by = "carrier") %>%
  mutate(age=year.x - year.y)

ggplot(join_carrier, aes(x=name, y=age, fill=name))+
geom_boxplot()+
theme_classic()+
labs(title = "Carrier Companies and Age of Planes", y = "Age Planes", x = "Carrier")+
  theme( axis.text.x = element_blank(),
         axis.ticks.x =element_blank(),
  )+
  scale_y_continuous(breaks =seq(0,40, by=10))

```
I see that most of the planes are fairly new between 1 to 15 years. It is interesting that among all carrier there is not a standard for replacing airplanes, which I see this in the distribution of the boxplot.

## Other graph 

```{r}


table2 <- select(chart_1, age,age_references,creation_year, dept_year, dep_delay,distance)
sum2 <- table2%>%
group_by(age_references)%>%
 summarise(tot_delays = sum(dep_delay > 0, na.rm = TRUE),
           distance = sum(distance> 0, na.rm = TRUE)
           ) 

ggplot(table2, aes(x = age, y = distance, fill=age_references)) +
  theme_light()+
  geom_bar(stat = "identity")+
  scale_y_continuous(breaks =seq(0,4e7, by=1e7))+
  labs(title = "Plane Age vs Distance", y = "Distance", x = "Plane Age")
  

```
The graph represent the planes by age and the flight distances. I would say that there are some old planes greater than 20 year with less distance flown. It may be that they were under utilize. It will be interesting to cross references with the cost of expenses.
