---
title: "Our World in Data Part 1s"
author: "Gabriel Guerrero"
date: "`r format(Sys.time(), '%B %d, %Y')`"
execute:
  keep-md: true
  warning: false
format:
  html:
    code-fold: true
    code-tools: true
---

## Background
The Our World in Data website publishes research and graphics that provide insight into world data. We are going to recreate one of their charts and use it to practice our annotation skills.

If the thing you want to edit is not a theme element, then it is related to data and can mostly likely be changed inside a geom_() or a scale_() layer.

## Load Packages

```{r}
#| label: load_packages
#| echo: false

# Use this R-Chunk to load the libraries you will use in this file.
library(tidyverse)
library(readr)

```
## Step1
Recreate this graph as closely as you can. This will/should take the bulk of your time
```{r}    
#| label: load
df <- read_csv("fruit.csv")
head(df,5)

df <- rename(df,fruit_pp = `Fruit | 00002919 || Food available for consumption | 0645pc || kilograms per year per capita`, 
         gdp_per_cap = `GDP per capita, PPP (constant 2017 international $)`,
         entity = Entity,
         code = Code,
         year = Year,
         continent = Continent)

# fill in missing values and do some filtering
data_2020 <- df |> 
  group_by(entity) |> 
    fill(continent, .direction = "downup") |> # fills in missing values in the continent column within each entity using the value(s) present in continent.
  ungroup() |> 
  filter(!is.na(continent)) |> # Non countries (continents and other country groups) don't have a continent value. Remove them.
  filter(year == 2020) # focus on year 2020 to match the visual on the webpage.

head(data_2020,5)
#colnames(data_2020)
#str(df)
```
## Graph
```{r}
#| label: graph
#| echo: false
#| #| fig-width: 10
#| fig-height: 3.5

graf <- ggplot(data_2020, aes(log(gdp_per_cap),fruit_pp))+
  geom_point(aes(colour= continent, size=entity), size=2.1)+
  theme_bw()+
  theme(panel.grid = element_line(color = "#e2e2e2",
                                  size = 0.5,
                                  linetype = "dashed"),
        panel.grid.minor = element_blank(),
        legend.title = element_blank()
        )+
  
  coord_cartesian(ylim = c(0, 370))+
  scale_x_continuous(breaks = log(c(1000, 2000, 5000, 10000, 20000, 50000, 100000)),
                     labels = c("$1,000", "$2,000", "$5,000", "$10,000", "$20,000","", "$100,000"), trans = "log")+
  
  scale_y_continuous(breaks =seq(0,350, by=50), 
                     labels = c("0 kg","50 kg", "100 kg", "150 kg", "200 kg","250 kg","300 kg", "350 kg"),
                     expand=expansion(add = c(0.4,0)))+
  
  scale_color_manual(values = c("#a2559c","#00847e", "#4c6a9c", "#e56e5a","#9a5129","#883039"))+
  guides(color = guide_legend( 
  override.aes=list(shape = 15, size=3)))

graf+ labs(title="Fruit consumption vs. GDP per capita, 2020",
         subtitle = "Average per capita fruit consumption, measured in kilograms per year versus gross domestic
product (GDP) per capita, measured in constant international-$.",
         caption = "Source: Food and Agriculture Organization of the United Nations, Data compiled from multiple sources by World Bank
OurWorldInData.org/diet-compositions • CC BY")+
    ylab("Fruit supply per person")+
    xlab("GDP per capita")+
   theme(
    plot.caption = element_text(color="black", hjust=0, size=9, lineheight = 1.2),
    plot.caption.position = "plot",
    plot.title.position ="plot",
    plot.title = element_text(face="bold", color="black", hjust=0, size=15, lineheight = 1.2),
    legend.text = element_text(size=12),
    legend.justification = "top",
    plot.subtitle = element_text(color="black", hjust=0, size=11, lineheight = 1),
    axis.text = element_text(size=11, lineheight = 2),
    axis.text.x = element_text(margin = margin(t = 0, r = 20, b = 2, l = 0),hjust = .6,lineheight = 1),
    axis.text.y = element_text(margin = margin(r = 0)),
    panel.border = element_blank(), axis.line = element_line(colour = "black"),
    axis.line.y = element_blank(),
    axis.ticks =element_blank()
    
   )

```
##Graph2

```{r}
#| label: graph_2
#| echo: false
#| #| fig-width: 10
#| fig-height: 3.5

garf1 <- filter(data_2020, entity=="Canada"| entity=="Mexico"| entity=="United States"| entity=="Ecuador")
  

graf1 <- ggplot(garf1, aes(log(gdp_per_cap),fruit_pp))+
  geom_area(data=data_2020,aes(log(gdp_per_cap),fruit_pp))+
  geom_col(aes(fill= entity))+
  theme_bw()+
  theme(panel.grid = element_line(color = "#e2e2e2",
                                  size = 0.5,
                                  linetype = "dashed"),
        panel.grid.minor = element_blank(),
        legend.title = element_blank()
        )+
  
  coord_cartesian(ylim = c(0, 370))+
  scale_x_continuous(breaks = log(c(1000, 2000, 5000, 10000, 20000, 50000, 100000)),
                     labels = c("$1,000", "$2,000", "$5,000", "$10,000", "$20,000","", "$100,000"), trans = "log")+
  
  scale_y_continuous(breaks =seq(0,350, by=50), 
                     labels = c("0 kg","50 kg", "100 kg", "150 kg", "200 kg","250 kg","300 kg", "350 kg"),
                     expand=expansion(add = c(0.4,0)))+
  
  scale_color_manual(values = c("#a2559c","#00847e", "#4c6a9c", "#e56e5a","#9a5129","#883039"))+
  guides(color = guide_legend( 
  override.aes=list(shape = 15, size=3)))


graf1+ labs(title="Selected Countries vs Fruit consumption vs. GDP per capita",
         subtitle = "Average per capita fruit consumption in USA, Mexico, Ecuador, Canada",
         caption = "Source: Food and Agriculture Organization of the United Nations, Data compiled from multiple sources by World Bank
OurWorldInData.org/diet-compositions • CC BY")+
    ylab("Fruit supply per person")+
    xlab("GDP per capita")+
   theme(
    plot.caption = element_text(color="black", hjust=0, size=9, lineheight = 1.2),
    plot.caption.position = "plot",
    plot.title.position ="plot",
    legend.justification = "center",
    plot.subtitle = element_text(color="black", hjust=0, size=11, lineheight = 1),
    panel.border = element_blank(), axis.line = element_line(colour = "black"),
    axis.line.y = element_blank(),
    axis.ticks =element_blank()
    
   )
```
### graph 3

```{r}
data3 <- data_2020 %>% 
  filter(continent %in% c('North America'))

ggplot(data = data_2020, aes(x=log(gdp_per_cap), y = fruit_pp)) +
  geom_point(alpha = 0.5) +
  geom_point(data = data3, aes(x = log(gdp_per_cap), y = fruit_pp, color = continent), alpha = 1, size = 3) +
  labs(y='Fruit supply per person', x = 'GDP per capita', title = 'Fruit consumption vs. GDP per capita, 2020', subtitle = 'Average per capita fruit consumption, measured in kilograms per year \nversus gross domestic product (GDP)', caption = 'Source: Food and Agriculture Organization of the United Nations, 
Data compiled from multiple sources by World Bank

OurWorldInData.org/diet-compositions • CC BY') 

```

## Summarizing
This is one of the longest projects I have worked. I have learn several commands such as:

   plot.caption = element_text()
    plot.caption.position = "plot"
    plot.title.position ="plot"
    plot.title = element_text(2),
    legend.text = element_text(),
    legend.justification = "top",
    plot.subtitle = element_text(),
    axis.text = element_text(),
    axis.text.x = element_text(),
    axis.text.y = element_text(),
    panel.border = element_blank(),
    axis.line.y = element_blank(),
    axis.ticks =element_blank()

Basically, I learned that there is a command to customize any part of the graph. When I do not know how google is a great tool to search for examples. 
Highlighting is also an important tool. I learned that manipulating the colors, shapes, and sizes is a great way to attract the reader´s eye. 
